A Discrepancy-based Framework to Compare Robustness between Multi-Attribute Evaluations

### 2016-12-11 Docs/Communication/discrepancy.tex
slides_v1

\sframe{Complex systems}{

@@

@@

\sframe{Multi-attribute Evaluation in Complex Socio-technical Systems}{

@@

$\rightarrow$

@@

.

@@

@@

$\rightarrow$ 
\medskip

$\rightarrow$

@@

\sframe{Theoretical Framing}{

@@

@@

@@

\textbf{Objectives as Spatial Integrals}

@@

@@

@@

@@

Territorial Systems $S_{i}=(\mathbf{X}_{i},\mathbf{Y}_{i})\in\mathcal{X}_{i}\times\mathcal{Y}_{i}$ with $\mathcal{X}_{i}=\prod_{k}\mathcal{X}_{i,k}$

\[
(\mathcal{X},\mathcal{Y}) \underset{def}{=} \left(\prod\tilde{\mathcal{X}}_{c}\right)\times\left(\prod\tilde{\mathcal{Y}}_{c}\right) = \left(\prod_{\mathcal{X}_{i,k}\in\mathcal{D}_{\mathcal{X}}}\mathbb{R}^{p_{i,k}^{X}}\right)\times\left(\prod_{\mathcal{Y}_{i,k}\in\mathcal{D}_{\mathcal{Y}}}\mathbb{R}^{p_{i,k}^{Y}}\right)
\]

Objectives : $H_{c}$ space of real-valued functions on $(\tilde{\mathcal{X}}_{c},\tilde{\mathcal{Y}}_{c})$, such that for all $h\in H_{c}$ :
\begin{enumerate}
\item $h$ is ``enough'' regular (tempered distributions e.g.)
\item $q_c=\int_{(\tilde{\mathcal{X}}_{c},\tilde{\mathcal{Y}}_{c})}h$ is a function describing the ``urban fact'' (the indicator in itself)
\end{enumerate}

@@

\sframe{Formal Description (III)}{

A relative \textit{Robustness Ratio} can thus be defined between two evaluations :

@@

%%%%%%%%%%%%%%%%%
\section{Results}
%%%%%%%%%%%%%%%%%

@@

@@

@@

Metropolitan Segregation on Ile-de-France, Insee income data (2011)

@@

\begin{itemize}
\item Spatial autocorrelation Moran index, defined as weighted normalized covariance of median income by $\rho = \frac{N}{\sum_{ij}w_{ij}}\cdot \frac{\sum_{ij}w_{ij}\left(X_i - \bar{X}\right)\left(X_j - \bar{X}\right)}{\sum_i \left(X_i - \bar{X}\right)^2}$
\item Dissimilarity index (close to Moran but integrating local dissimilarities rather than correlations), given by $d =  \frac{1}{\sum_{ij}w_{ij}} \sum_{ij} w_{ij} \left|\tilde{X}_i - \tilde{X}_j\right|$\\ with $\tilde{X}_i = \frac{X_i - \min(X_k)}{\max(X_k) - \min(X_k)}$
\item Complementary of the entropy of income distribution that is a way to capture global inequalities $\varepsilon = 1 + \frac{1}{\log(N)} \sum_i \frac{X_i}{\sum_k X_k} \cdot \log\left(\frac{X_i}{\sum_k X_k}\right)$
\end{itemize}

@@

@@

\textit{Example of Segregation maps}

@@

% missing data
\textit{Framework Application : sensitivity to missing data}

@@

\includegraphics[width=0.45\textwidth]{figures/alldeps_rob_renormindics}

\includegraphics[width=0.45\textwidth]{figures/alldeps_robsd_renormindics}

@@

%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%

\sframe{Applicability}{

$\rightarrow$ Application to decision-making procedures : adding robustness as a dimension ?

@@

@@

$\rightarrow$ Assumptions validity ranges : some indicators may difficultly be viewed as spatial integrals (as some accessibility measures~\cite{kwan1998space}

@@

\sframe{Further Developments}{

@@

\sframe{Conclusion}{

$\rightarrow$

\bigskip
\bigskip
\bigskip

\footnotesize{ - All code available at \texttt{https://github.com/JusteRaimbault/RobustnessDiscrepancy}

 - Paper preprint available at \texttt{http://arxiv.org/abs/}
}

@@

>>>>

\subsection{Context}

\sframe{Multi-objective Complex Systems}{

@@

\textit{From morphogenetic factors to optimal design, fundamental multi-objective nature of optimization processes}

\includegraphics[width=0.35\textwidth]{figures/lattice}\hspace{0.5cm}
\includegraphics[width=0.55\textwidth]{figures/exampleOptimalCorridor}

\smallskip

{\footnotesize Source : Left \cite{raimbault2014hybrid} ; Right \cite{raimbault2015labex}}

@@

\sframe{Multi-objective Evaluations of Socio-technical Systems}{

@@

\bigskip
\bigskip

$\rightarrow$ Design and understanding (i.e. evaluation) of Socio-technical systems at the intersection as hybrid systems~\cite{picon2013smart}\cite{haken2003face}

@@

\textit{Reliability of an evaluation is crucial; various approaches to robustness depending on method}

\bigskip
\bigskip

$\rightarrow$ Naturally included in the construction and estimation of statistical models~\cite{launer2014robustness} (e.g. p-value, beta power, AIC)

%\bigskip

%$\rightarrow$ Bayesian statistics also include robustness of sequential estimation

$\rightarrow$ In multi-objective optimization, diverse methods: sensitivity of Pareto front to perturbations~\cite{deb2006introducing};  continuity of solutions

\cite{1688537}

$\rightarrow$ \cite{dobbie2013robustness} studies robustness for multi-attribute evaluations as biases depending on weighting techniques

@@

\subsection{Research Question}

@@

% -> why multi -attribute
% -> data-driven
% -> model-independant

\bigskip
\bigskip

$\rightarrow$ Model-independence and method-independence; framework based only on data structure (and thus quality) and indicator values
\medskip

$\rightarrow$ Particular case of multi-attribute evaluations, where dimensions are aggregated, in order to obtain a simple measure of robustness

@@

\subsection{Theoretical Framing}

\sframe{Intuitive and Theoretical Framing}{

@@

\begin{enumerate}
\item {\justify Systems are seen from the perspective of raw data available: data-driven approach}

\smallskip

\item  A choice of indicators captures the realization of an ``urban fact''~\cite{mangin1999projet}, in the sense of stylized process with different spatial realizations

\smallskip

\item Given many systems and indicators, a common space can be build to compare them

\smallskip

\item Discrepancy of data~\cite{dick2010digital} in that space captures various aspects linked to robustness: system scale and range, precision of data, missing data

\smallskip

\item Robustness must also capture an error done on indicator computation

\end{enumerate}

@@

\subsection{Assumptions}

@@

\textbf{Objectives as Kernel Integrals}

@@

$\rightarrow$ Kernel functions exist for each objective, computed as their integrals

$\rightarrow$ Reasonable for most systems, as e.g. for territorial systems  it is analog to smoothing of Geographically Weighted Regression~\cite{brunsdon1998geographically}

@@

$\rightarrow$ Aggregated objective as linear combination of attributes $q(\vec{x})=\sum_i{w_i q_i(\vec{x})}$

$\rightarrow$ Choice of weights at the core of decision-making process; not our scope here, we take simply relative indicator importance $w_{i,c}^{L}=\frac{\hat{q}_{i,c}}{\sum_{c}\hat{q}_{i,c}}$

@@

\subsection{Formal Description}

@@

\textbf{Territorial Systems : } Data $S_{i}=\mathbf{X}_{i}\in\mathcal{X}_{i}$ with $\mathcal{X}_{i}=\prod_{k}\mathcal{X}_{i,k}$

\textbf{System space}

\[
\mathcal{X} \underset{def}{=} \left(\prod\tilde{\mathcal{X}}_{c}\right) = \left(\prod_{\mathcal{X}_{i,k}\in\mathcal{D}_{\mathcal{X}}}\mathbb{R}^{p_{i,k}^{X}}\right)
\]

\textbf{Objectives :} $H_{c}$ space of real-valued functions on $\tilde{\mathcal{X}}_{c}$, such that :
\begin{enumerate}
\item $h_c\in H_{c}$ are ``enough'' regular (tempered distributions e.g.)
\item $q_c=\int_{\tilde{\mathcal{X}}_{c}}h$ is a function describing the ``urban fact'' (the indicator in itself)
\item Normalized kernels $h_c (\vec{x}) \in [0,1]$ % TYPO PAPER HERE ? not clear beween h_c and q_c ; both can be and are in applications
\end{enumerate}

@@

$\rightarrow$ A relative \textit{Robustness Ratio} can thus be defined between two evaluations :

@@

%%%%%%%%%%%%%%%%%
\section{Application}
%%%%%%%%%%%%%%%%%

\subsection{Synthetic Data}

@@

\textit{Using OpenStreetMap roads and buildings data, construction of synthetic indicators on Paris districts (car daily use, car flows in streets, relative length of pedestrian streets), computed in Monte-carlo simulations}

$\rightarrow$ Minimal ratio (relative to 1st Arr.) in 15th Arr. ($0.92 \pm 0.03$), intuitively expected

\centering

\includegraphics[width=0.5\textwidth]{figures/ex_data}

\smallskip

\small \textit{Example of raw data}

@@

\subsection{Metropolitan Segregation}

@@

\textit{Application to metropolitan Segregation on Ile-de-France, Insee income data (2011)}

\bigskip
\bigskip

@@

\begin{itemize}
\item Spatial autocorrelation Moran index
%defined as weighted normalized covariance of median income by $\rho = \frac{N}{\sum_{ij}w_{ij}}\cdot \frac{\sum_{ij}w_{ij}\left(X_i - \bar{X}\right)\left(X_j - \bar{X}\right)}{\sum_i \left(X_i - \bar{X}\right)^2}$
\item Dissimilarity index $d =  \frac{1}{\sum_{ij}w_{ij}} \sum_{ij} w_{ij} \left|\tilde{X}_i - \tilde{X}_j\right|$
%with $\tilde{X}_i = \frac{X_i - \min(X_k)}{\max(X_k) - \min(X_k)}$
\item Complementary of distribution entropy $\varepsilon = 1 + \frac{1}{\log(N)} \sum_i \frac{X_i}{\sum_k X_k} \cdot \log\left(\frac{X_i}{\sum_k X_k}\right)$
\end{itemize}

@@

\sframe{Metropolitan Segregation: Indicators}{

@@

\textit{Example of Segregation maps (linking median income with local Moran index)}

@@

\sframe{Metropolitan Segregation: Results}{

% missing data
\textit{Framework Application : robustness comparison between administrative areas and sensitivity to missing data}

\centering

\includegraphics[width=0.55\textwidth]{figures/alldeps_rob_renormindics}
\includegraphics[width=0.55\textwidth]{figures/alldeps_robsd_renormindics}

\small \textit{(Left) Robustness ratio compared to full region for each area, as a function of missing data proportion; (Right) Standard deviations}

\subsection{Discussion}

\sframe{Discussion}{

\vspace{-0.5cm}

\textbf{Applicability}

@@

$\rightarrow$ Application to decision-making procedures : adding robustness as a dimension

\smallskip

$\rightarrow$ Assumptions validity ranges : some indicators may difficultly be viewed as spatial integrals (as some accessibility measures~\cite{kwan1998space})

\smallskip

\textbf{Further Developments}

\smallskip

\smallskip

$\rightarrow$ Multi-dimensional measure of robustness using discrepancy along different dimensions

@@

\sframe{Conclusion}{
\justify

$\rightarrow$ An original data-driven approach to robustness in multi-attribute evaluations of complex socio-technical systems; independent of model and method

@@

$\rightarrow$ Future insights from application to diverse system types and fields should refine the framework

@@

$\rightarrow$ Importance of interdisciplinarity (linking here statistics with computational modeling), crucial to study Complex Systems

\footnotesize{

- Special thanks to J. Keutchayan (Ecole Polytechnique de Montr{\'e}al) for suggesting the original idea of using discrepancy

- All code and data available at \texttt{https://github.com/JusteRaimbault/RobustnessDiscrepancy}

 - Paper preprint available \cite{raimbault2016discrepancy} at \texttt{http://arxiv.org/abs/1608.00840}
}

@@

\sframe{Reserve Slides}{

\centering

\Large \textbf{Reserve Slides}

@@

\sframe{Kernel Examples}{

Typical concrete example of kernels can be :

\begin{itemize}
\item A mean of rows of $\mathbf{X}_{i,c}$ is computed with $h(x)=x\cdot f_{i,c}(x)$ where $f_{i,c}$ is the density of the distribution of the assumed underlying variable.
\item A rate of elements respecting a given condition $C$, $h(x)=f_{i,c}(x)\chi_{C(x)}$ 
\item For already aggregated variables $\mathbf{Y}$, a Dirac distribution allows to express them also as a kernel integral. 
\end{itemize}

@@

\sframe{Implementation}{

\begin{itemize}
\item Preprocessing of geographical data is made through QGIS

\cite{qgis2011quantum}
\item Core implementation of the framework is done in R \cite{team2000r} for the flexibility of data management and statistical computations 
\item package \texttt{DiceDesign}\cite{franco20092} for computation of discrepancies
\end{itemize}

\sframe{Synthetic Indicators (I)}{
\begin{itemize}
\item Complementary of the average daily distance to work with car per individual, approximated by, with $n_{cars}(b)$ number of cars in the building (randomly generated by associated of cars to a number of building proportional to motorization rate $\alpha_m ~ 0.4$ in Paris), $d_w$ distance to work of individuals (generated from the building to a uniformly generated random point in spatial extent of the dataset), and $d_{max}$ the diameter of Paris area, $\bar{d}_w = 1 - \frac{1}{|b\in A(a)|} \cdot \sum_{b\in A(a)}{n_{cars}(b)\cdot \frac{d_w}{d_{max}}}$

\end{itemize}

\sframe{Synthetic Indicators (II)}{

\begin{itemize}
\item Complementary of average car flows within the streets in the district, approximated by, with $\varphi(s)$ relative flow in street segment $s$, generated through the minimum of 1 and a log-normal distribution adjusted to have $95\%$ of mass smaller than 1 what mimics the hierarchical distribution of street use (corresponding to betweenness centrality), and $l(s)$ segment length, $\bar{\varphi} = 1 - \frac{1}{|s\in A(a)|} \cdot \sum_{s \in A(a)}{\varphi(s)\cdot \frac{l(s)}{\max{(l(s))}}}$
\item Relative length of pedestrian streets $\bar{p}$, computed through a randomly uniformly generated dummy variable adjusted to have a fixed global proportion of segments that are pedestrian.
\end{itemize}
}

\begin{frame}
\frametitle{Synthetic Numerical Results}

\includegraphics[width=\textwidth,height=0.9\textheight]{figures/syntheticnum.png}

\end{frame}
